\documentclass{article}
\usepackage{amsmath, amssymb, amsfonts, amsthm, mathtools}
\usepackage[utf8]{inputenc}
\usepackage[inline]{enumitem}
\usepackage{cancel}
\usepackage{soul}
\usepackage{hyperref}
\newtheorem{theorem}{Theorem}
\newtheorem{lem}{Lemma}
\newtheorem{defn}{Definition}

\setlength\parindent{0pt}

\usepackage{xcolor}
\definecolor{mybgcolor}{RGB}{46, 51, 63} %46, 51, 63

\usepackage{pagecolor}
\pagecolor{mybgcolor}
\color{white}

\usepackage{geometry}
\geometry{
	a4paper,
	total={170mm,257mm},
	left=20mm,
	top=20mm,
}

\title{Short Quiz 4: Solution}      % here
\author{Aryaman Maithani}
\date{28th August, 2019}			 % here

\begin{document}
\maketitle

\hrulefill

\textbf{Question.}
Recall that a cubic polynomial is a function $q:\mathbb{R} \to \mathbb{R}$ given by $q(x) = ax^3 + bx^2 + cx + d$ for $a,\;b,\;c,\;d \in \mathbb{R}$ and $a \neq 0.$\\
State whether the following statement is true or false. Justify your answer.\\ 
Every cubic polynomial has an inflection point.
\hfill [5 marks]
\begin{flushright}
	[2 marks for correct alternative (T/F); 3 marks for correct justification]\\~\\
\end{flushright}

\hrulefill

\textbf{Answer.} T \hfill \textbf{[2]}\\  % Fill T/F
Justification: \\
We will justify our answer using the third derivative test. \\
First note that a cubic polynomial is thrice differentiable. \hfill \textbf{[1]}\\
We compute the following derivatives: $f''(x) = 6ax + b$ and $f'''(x) = 6a.$ \hfill \textbf{[1]}\\
As $a \neq 0,$ the following quantity is well-defined, $x_0 := -\frac{b}{3a}.$ Moreover, $f''(x_0) = 0$ and $f'''(x_0) = 6a \neq 0.$ \hfill \textbf{[1]}\\
This is a \emph{sufficient} condition for an inflection point. Thus, we have shown that every cubic polynomial has an inflection point.

\hrulefill

\vspace{0.2 cm}

Points to be noted -
\begin{enumerate}
	\item We do have an alternate solution which relies only on the function being twice differentiable and checking the sign of $f''$ around $-\frac{b}{3a}.$
	\item Half a mark has been deducted for those who simply wrote that $f''(x) > 0$ for $x > -\frac{b}{3a}$ as this is true only when $a > 0.$ 
	\item Marks have been appropriately deducted when a student has written a wrong statement, some common ones being - 
	\begin{enumerate}[nosep] 
		\item $f''(c) = 0$ is a sufficient condition for an inflection point.\\
		Counterexample - $f(x) = x^4$ and $c = 0.$
		\item $f''(c) = 0$ and $f'''(c) \neq 0$ is a necessary condition for an inflection point.\\
		Counterexample - $f(x) = x^5$ and $c = 0.$\\
		This is especially wrong as is not even necessary for a function to be continuous at a point of inflection.
	\end{enumerate}
\end{enumerate}

\end{document}