\documentclass{article}
\usepackage{amsmath, amssymb, amsfonts, amsthm, mathtools}
\usepackage[utf8]{inputenc}
\usepackage[inline]{enumitem}
\usepackage{cancel}
\usepackage{soul}
\usepackage{hyperref}
\newtheorem{theorem}{Theorem}
\newtheorem{lem}{Lemma}
\newtheorem{defn}{Definition}

\setlength\parindent{0pt}

\usepackage{xcolor}
\definecolor{mybgcolor}{RGB}{46, 51, 63} %46, 51, 63

\usepackage{pagecolor}
\pagecolor{mybgcolor}
\color{white}

\usepackage{geometry}
\geometry{
	a4paper,
	total={170mm,257mm},
	left=20mm,
	top=20mm,
}

\title{Short Quiz 3: Solution}      % here
\author{Aryaman Maithani}
\date{21st August, 2019}			 % here

\begin{document}
\maketitle

\hrulefill

\textbf{Question.} State whether the following statement is true or false. Justify your answer.\\
Is it possible to find a function $f : \mathbb{R} \to \mathbb{R}$ such that
\begin{enumerate}[label = (\alph*)]
	\item $f$ is continuous on $[-7, 0]$
	\item $f$ is differentiable on $(-5, 0)$
	\item $f'(x) \le 2$
	\item $f(-5) = -3$ and $f(0) = 8.$
\end{enumerate}
\begin{flushright}
	[2 marks for correct alternative (T/F); 3 marks for correct justification]\\~\\
\end{flushright}

\hrulefill

\textbf{Answer.} F \hfill \textbf{[2]}\\  % Fill T/F
Justification: \\
Assume that such a function $f$ exists.\\
By hypothesis, $f$ is continuous on $[-7, 0];$ thus, it is continuous on $[-5, 0].$ \hfill \textbf{[0.5]}\\
Moreover, $f$ is differentiable on $(-5, 0).$ \hfill \textbf{[0.5]}\\
Thus, by MVT, we know that there exists $c \in (-5, 0)$ such that
\[f'(c) = \dfrac{f(0) - f(-5)}{0 - (-5)} = \dfrac{11}{5} > 2\]
\begin{flushright}
	\textbf{[2]}
\end{flushright}	
However, by hypothesis $f'(c) \le 2.$ Thus, we have arrived at a contradiction. \hfill $\blacksquare$\\

\hrulefill

\vspace{0.2 cm}

Points to be noted -
\begin{enumerate} 
	\item Those who have not mentioned that $f$ is continuous on $[-5, 0]$ have lost half a mark as per the above scheme. It is important to mention that as we cannot appeal to the Mean Value Theorem otherwise.
	\item For those who written that $11/5 = 2.1$ or $11/5 = 2.5,$ a quarter mark has been deducted. Note that it was not necessary to calculate $11/5$ in its decimal expansion. However, if you write something that is incorrect and use that to justify your claim, it is not be correct.
	\item Half a mark has been deducted for simply writing $f'(c) = \dfrac{f(0) - f(-5)}{0 - (-5)}$ without mentioning what $c$ is, that is, not writing ``for some $c$ in $(-5, 0)$'' or ``$\exists c \in (-5, 0).$''
\end{enumerate}

\end{document}